---
title: "Counting"
author: "Ho Ngoc Van"
date: "3/1/2021"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

**Question 1:** How many ways are there to permute the letters in the word MISSISSIPPI?

**Answer:** $\color{red}{\text{$\frac{11!}{1!4!4!2!}$}}$

**Code:**

Small example

```{r}
library(randtests)
s <- c("M", "I", "M", "P")
unique(permut(s))
factorial(length(s)) / prod(factorial(table(s)))
```

Result

```{r}
s <- c("M", "I", "S", "S", "I", "S", "S", "I", "P", "P", "I")

factorial(length(s)) / prod(factorial(table(s)))
```

---

**Question 2:**

a.How many 7-digit phone numbers are possible, assuming that the first digit can't be a 0 or a 1?

**Answer:** $\color{red}{\text{$8*10^6$}}$

**Code:**

```{r}
8*10^6
```

b.Re-solve (a), expect now assume also that the phone number is not allowed to start with 911?

**Answer:** $\color{red}{\text{$7*9*9*10^4$}}$

**Code:**

```{r}
7*9*9*10^4
```

---

**Question 3:** Fred is planning to go out to dinner each night of a certain week, Monday through Friday, with each dinner being at one of his ten favorite restaurants.

a.How many possibilities are there for Fred's schedule of dinner for that Monday through Friday. If Fred is not willing to eat at the same restaurant more than once?

**Answer:** $\color{red}{\text{$10*9*8*7*6$}}$

**Code:**

```{r}
10*9*8*7*6
```

b.If Fred is willing to eat at the same restaurant more than once, but is not willing to eat at the same place twice in a row (or more)?

**Answer:** $\color{red}{\text{$10*9*9*9*9$}}$

**Code:**

```{r}
10*9*9*9*9
```

---

**Question 4:** A *round-robin tournament* is being held with *n* tennis players; this means that every player will play against every other player exactly once

a.How many possible outcomes are there for the tournament (the outcome lists out who won and who lost for each game)?

**Answer:** $\color{red}{\text{$2^\binom n2$}}$

**Code:**

```{r}
comb = function(n, x) {
  factorial(n) / factorial(n-x) / factorial(x)
}

n <- 7
2^comb(n, 2)
```

b.How many games are played in total?

**Answer:** $\color{red}{\text{$\binom n2$}}$

**Code:**

```{r}
comb(n, 2)
```

---

**Question 5:** A *knock-out tournament* is being held with $2^n$ tennis players. This means that for each round, the winners move on to the next round and the losers are eliminated, until only one person remains. For example, if initially there are $2^4=16$ players, then there are 8 games in the first round, then the 8 winners move on to round 2, then the 4 winners move on to round 3, then the 2 winners move on to round 4, the winner of which is declared the winner of the tournament. (There are various systems for determining who plays whom within a round, but these do not matter for this problem.)

a.How many rounds are there?

**Answer:** $\color{red}{\text{n}}$

b.Count how many games in total are played, by adding up the numbers of games played in each round?

**Answer:** $\color{red}{\text{$\sum_{\substack{1<=k<=n}}2^{k-1}=2^n-1$}}$

**Code:**

```{r}
curve(2^x-1, -5, 5, xlab = "n", ylab = "2^n-1")
```

c.Count how many games in total are played, this time by directly thinking about it without doing almost any calculation?

**Hint: ** $\color{green}{\text{How many players need to be eliminated?}}$

**Answer:** $\color{red}{\text{#players are eliminated = 2^n - 1}}$

---

**Question 6:** There are 20 people at a chess club on a certain day. They each find opponents and start playing. How mnay possibilities are there for how they are matched up, assuming that in each game it does matter who has the white pieces (in a chess game, one player has the white pieces and the other player has the black pieces)

**Answer:** $\color{red}{\text{$2*\binom {20}2$}}$

**Code:**

```{r}
comb = function(n, x) {
  factorial(n) / factorial(n-x) / factorial(x)
}

2*comb(20, 2)
```

---

**Question 7:** 

**Answer:** $\color{red}{\text{fill}}$

**Code:**

```{r}

```

---

**Question 8:** 

**Answer:** $\color{red}{\text{fill}}$

**Code:**

```{r}

```

---

**Question 9:** 

**Answer:** $\color{red}{\text{fill}}$

**Code:**

```{r}

```

---

**Question 10:** 

**Answer:** $\color{red}{\text{fill}}$

**Code:**

```{r}

```

---

**Question 11:** 

**Answer:** $\color{red}{\text{fill}}$

**Code:**

```{r}

```

---

**Question 12:** 

**Answer:** $\color{red}{\text{fill}}$

**Code:**

```{r}

```

---

**Question 13:** 

**Answer:** $\color{red}{\text{fill}}$

**Code:**

```{r}

```

---

**Question 14:** 

**Answer:** $\color{red}{\text{fill}}$

**Code:**

```{r}

```

---
